---
title: "lab5_code_along"
author: "Alyssa Kibbe"
date: "11/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
library(effsize)
library(broom)
library(ggbeeswarm)
library(car)
```

# Part 1. t-test
### do we have enough evidence to show that they are from populations of different means
Mass comparision: male & female chinstrap penguins

```{r}
chinstrap <- penguins %>% 
  filter(species == "Chinstrap")

# histograms
ggplot(data = chinstrap, aes(x = body_mass_g))+
  geom_histogram(bins = 10) +
  facet_wrap(~sex)

#QQ-plot
ggplot(data = chinstrap, aes(sample = body_mass_g))+
  geom_qq()+
  facet_wrap(~sex)

# distriptive stastistics
chinstrap %>% 
  group_by(sex) %>% 
  summarize(
    mean_mass = mean(body_mass_g),
    median_mass = median(body_mass_g),
    sd_mass = sd(body_mass_g),
    sample_size = n()
  )


```
##### look mostly normal, mean would be reasonable. QQ plot shows they are very straight which means that its pretty normal. 

## now to compare sizes
# Get vectors of body_mass_g for female and male chinstrap penguins separately. 
```{r}
chinstrap_f <- chinstrap %>% 
  filter(sex == "female") %>% 
  pull(body_mass_g)

chinstrap_m <- chinstrap %>% 
  filter(sex == "male") %>% 
  pull(body_mass_g)
```

### run t-test in a couple diff ways

1. is there a significant difference in body mass between male and female chinstrap penguins? (two-sided)

```{r}
chinstrap_ttest <- t.test(chinstrap_f, chinstrap_m)
chinstrap_ttest

cohen.d(chinstrap_f, chinstrap_m)

```
### tiny p-value means that there is a very tiny chance that means are from the same mean populations. there is a significant difference between male and female chinstrap penguins.

#### cohen's d shows large effect size

are male chinstrap penguins significantly larger (body mass) than females (one-tailed)?

```{r}
t.test(chinstrap_m, chinstrap_f, alternative = "greater")
```






